---
#----ESSENTIALS----#
title: '\fontfamily{phv}\selectfont Predicting outcome in prostate cancer patients using a multi-signature risk classifier, derived from urinary extracellular vesicles'
author: "\\large\\textbf{Shea P. Connell}\\normalsize^1^, Marcel Hanna^1^, Frank McCarthy^2^, Rachel Hurst^1^, Helen Curley^1^, Martyn Webb^1^, Helen Walker^3^, Rob Mills^3^, Richard Y. Ball^3^, Martin G. Sanda^4^, Kathryn L. Pellegrini^4^, Dattatraya Patil^4^, Antoinette S. Perry^5^, Jack Schalken^6^, Hardev Pandha^7^, Hayley Whitaker^8^, Nening Dennis^2^, Christine Stuttle^2^, Ian G. Mills^9,10,11^, Ingrid Guldvik^10^, The Movember GAP1 Urine Biomarker Consortium, Chris Parker^2,12^, Jeremy Clark^1^, Daniel S. Brewer^1,13^, Colin S. Cooper^1^"
affiliation: "^1^ Norwich Medical School, University of East Anglia, UK; ^2^ The Institute of Cancer Research, UK; ^3^ Norfolk and Norwich University Hospitals NHS Foundation Trust, UK; ^4^ Department of Urology, Winship Cancer Institute, Emory University School of Medicine, USA; ^5^ Cancer Biology and Therapeutics Laboratory, School of Biology and Environmental Science, Conway Institute, University College Dublin, Ireland; ^6^ Nijmegen Medical Centre, Radboud University Medical Centre, The Netherlands; ^7^ Faculty of Health and Medical Sciences, The University of Surrey, UK; ^8^ Molecular Diagnostics and Therapeutics Group, University College London, UK; ^9^ School of Medicine, Dentistry and Biomedical Sciences, Institute for Health Sciences, Centre for Cancer Research and Cell Biology, Queen's University Belfast, UK; ^10^ Centre for Molecular Medicine, University of Oslo, Norway; ^11^ Nuffield Department of Surgical Sciences, University of Oxford, UK; ^12^ The Royal Marsden Hospital, Sutton, UK; ^13^ The Earlham Institute, Norwich Research Park, UK"
#----FONT STYLE----#
base_font_size: "30pt" #sets '\normalsize' (sizes of the title and sections headings will adjust based on it)
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
bibliography_size: "tiny"

#----COLOURS----#

#Background/ Line Colours
title_bgcol: "690F8C"  #Hex colour for background of poster title section (default is blue)
poster_bgcol: "ffffff" #colour of the poster main background
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "ffffff" #colour of other links within the poster
columnline_col: "000000" #colour 
header_bgcol: "008080"

#Text Colours
title_textcol: "ffffff" #colour of title text
author_textcol: "E5E5E5" # Colour of author text
affiliation_textcol: "B6B6B6" # Colour of affiliation text
body_textcol: "000000" # Colour of main text in the body of poster
header_textcol: "690F8C" # Colour of the poster header titles
footnote_textcol: "000000" # Colour of footnote text if used

#----POSTER SIZE & COLUMN FORMATTING----#
poster_height: "140cm" # height in inches of poster
poster_width: "90cm" # width in inches of poster
column_numbers: 2 # Number of columns that the poster has
column_margins: "50mm" # Margin spacing for columns

#----OTHER----#
bibliography: MyLibrary # name of the .bib file used for referencing
output: posterdown::posterdown_pdf
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis',fig.height = 3.5, fig.width = 4.5, out.width = "100%", warning = FALSE, fig.align = 'center', message = FALSE, dev='cairo_pdf')
library(readr)
library(kableExtra)
library(survival)
library(survminer)
library(pROC)
```
<!--- Define a bunch of colours to be used in the Quick Read box -->
\begin{figure}
\definecolor{bordercolour}{HTML}{DCDCDC}
\definecolor{titleColour}{HTML}{252525}
\definecolor{myframecolour}{HTML}{690F8C}
\mdfsetup{frametitlealignment=\center,
          frametitlefont=\fontfamily{ppl}\selectfont,
          frametitlefont=\huge,
          frametitlebackgroundcolor=bordercolour,
          frametitlefontcolor = titleColour,
          linecolor=black,
          linewidth=5pt}
\begin{mdframed}[backgroundcolor = myframecolour,
                  font = \fontfamily{ppl}\selectfont,
                  fontcolor = white, 
                  roundcorner = 20pt, 
                  innerleftmargin = 30pt,
                  innerrightmargin = 30pt,
                  frametitle={\textbf{Quick Read}}]

\Large\textbf{Question: }\large Can a non-invasive single-sample urine test reveal \textbf{both diagnostic \& prognostic information about 537 prostate cancer patients}, utilising extracellular vesicle-derived RNA expression patterns?\bigskip\\
\Large\textbf{Findings: }\large A robust, four-risk-signature model identified two groups with differing rates of treatment intervention in active surveillance use (\textbf{High Risk HR = 3.7, Low Risk HR = -7.0}) \& predicted initial biopsy outcome (\textbf{AUC = 0.81})\bigskip\\
\Large\textbf{Impact: }\large Clinical implementation of this model has the potential to \textbf{avoid the unnecessary initial biopsy of men} \& the repeated, invasive follow-up of men on active surveillance with indolent prostate cancer could be drastically reduced, or \textbf{provide a means for exiting surveillance altogether}. 
\end{mdframed}
\end{figure}
\vspace{-45mm}

# Introduction
\vspace{-7.5mm}
**Clinical problem:** To combat the over-treatment of men with indolent prostate cancer, men with D'Amico Low & favourable Intermediate risk cancers are often offered active surveillance (AS) as an option^\cite{Selvadurai2013}^. Involving regular assessment for signs of disease progression, AS has proven successful in reducing unnecessary treatment.But invasive and repeated assessment means elective treatment rates can be higher than 30%^\cite{Hamdy2016}^ and there is no formal method for patients to exit  AS. 

**Rationale:** Previous studies established urine as a suitable medium for non-invasive sampling of the prostate^\cite{McKiernan2016b,Tomlins2016a,Donovan2015,VanNeste2016}^. In this vein, we have developed a risk prediction model, using NanoString quantified RNA expression from urinary extracellular vesicles, with the aim to discriminate accurately between men with, and without, clinically significant cancers. 
\vspace{-28.5mm} 

# The Movember Cohort
\vspace{-7.5mm}
  **Patients \& Samples:** Post-DRE urine from 537 men with and without histologically proven cancer with a PSA < 100 ng/mL, collected from urology clinics in the UK, Ireland and USA between 2009 and 2015. Extracellular vesicle RNA profiles derived with NanoString. Biopsy information from trans-rectal ultrasound guided needle biopsy.  

  **AS Eligibility \& Progression:** Age 50–80, clinical stage T1/T2, PSA < 15 ng/mL, Gs <7 (Gs < 4+3 if age >65), and <50% percent positive biopsy cores. Progression was defined as PSA velocity >1 ng/mL per year or adverse histology on repeat biopsy, defined as primary G >3 or >50% biopsy cores positive for cancer. 

\vspace{-28.5mm}

# Methods
\vspace{-7.5mm}
**Model:** Continuation-ratio LASSO regression trained on 67% of the data \& treating D'Amico risk status as an ordinal variable. Four risk signatures generated predicted probabilities of normal tissue (R1), D’Amico Low-risk (R2), Intermediate-risk (R3), and High-risk (R4) PCa in a given sample. 

**Testing:**  ROC-AUC analysis tested prediction of biopsy outcome in the remaining 33% test dataset. Survival analyses assessed prognostication of disease progression of patient in an AS sub-cohort with long follow-up(n = 87, >5 years). 
\vspace{-28.5mm}

# Results
\vspace{-7.5mm}
**Prediction of D'Amico Risk \& biopsy outcome**: There was a strong association with clinical category, appearing to recategorise approximately 15% of samples at either end of each D'Amico risk group (Figure 1). AUC of a clinically significant initial biopsy outcome of D'Amico Intermediate- or High-Risk: **0.813 (95% CI: 0.779 - 0.847, Figure 2)**. 

**AS Prognostication** Two groups, dichotomised by a robust R4 threshold had large differences in time to progression; 60% of those deemed high risk progressed whilst only 10% of low risk patients progressed 5 years from urine sample collection: (Figure 3, median progression 26%). Cox models detailed significant differences in progression; **R4 HR = 3.71 (95% CI: 1.53 to 5.89)**, and **R1 HR =-7.03 (95% CI: -12.29 to -1.77)**. 
\vspace{-28.5mm}

# Conclusion
\vspace{-7.5mm}
Clinical implementation of our model for biopsy prediction could reduce the rate of unnecessary initial biopsy, performing similarly to previously published urine tests for clinically significant cancers. However, no other test currently exists to prognosticate AS patients as we have shown. The model has the potential identify men at particular risk of progression, or, more importantly, those who may be safely left alone for up to five years.  
```{r, fig.cap='ROC-AUC plot of the ability of R4 to predict the presence of D\'Amico Intermediate or High risk cancer on initial biopsy. Shaded region indicates 95\\% confidence intervals from 2000 bootstrap resamples.', cache=TRUE, fig.asp = 0.8, fig.width=5, out.extra='trim={0 5mm 0 10mm},clip'}
#Read in the pre-cleaned data:
Results <- readRDS("data/Results.RDS") 
par(bty = "n")
#Create the ROC object for R4:
R4roc <- roc(response = Results$ClinSig,
             predictor = Results$R4,
             ci = TRUE, 
             ci.method = "bootstrap",
             # arguments for plot
             plot=TRUE,
             print.auc=TRUE, 
             show.thres=TRUE,
             print.auc.x = 0.5,
             print.auc.y = 0.05,
             print.auc.cex = 0.85,
             asp = 0.75,
             cex.axis = 0.75,
             xlab = "",
             ylab = "",
             identity.col = "black"
             
)
# Compute and plot the 95% CI:
sens.ci <- ci.se(R4roc,
                 specificities = seq(0, 1, 0.05))
RedCol <- adjustcolor("red2", alpha.f = 0.4) #Custom colour allows for alpha
plot(sens.ci, 
     type ="shape",
     lwd = 0.1,
     col = RedCol)
# need to re-draw R4roc over the CI
plot(R4roc, add=TRUE, col = "red2")
mtext(side=1, line=3, "Specificity")
mtext(side=2, line=2, "Sensitivity")
```
\vspace{-10mm}

```{r, fig.asp=0.75, fig.cap='Kaplan-Meier plot of time to disease progression in months from initial urine collection. Colours indicate the dichotomised model thresholds, Green – Low R4, Red – High R4. Numbers above the x-axis indicate the number of patients in each group at risk at the given time intervals. Shaded regions indicate 95\\% confidence intervals.', out.extra='trim={0 2.5mm 0 2.5mm},clip'}
#Read in the pre-cleaned data:
ASResults <- readRDS("data/ASResults.RDS") 
#Optimise the threshold to find the best binary threshold that keeps at least 30% of patients in the smallest group
OptimalThresh <- surv_categorize(surv_cutpoint(data = ASResults,
                                                    time = "FUP",
                                                    event = "Progressed",
                                                    variables = "R1",
                                                    minprop = 0.30),
                                      labels = c("Low R1", "High R1"))
#Build the survival object 
SurvOptim <- survfit(Surv(FUP, Progressed) ~ R1, data = OptimalThresh)

#Construct the plot (using SurvMiner)
ggsurvplot(SurvOptim, 
           risk.table = TRUE,
           palette =  c("ForestGreen", "Red2"),
           legend = "none",
           risk.table.title =  "",
           risk.table.pos = "in",
           xlab = "Time (months)",
           ylab = "Proportion free of progression",
           tables.height = 0.25,
           font.x = c(8, "bold"),
           font.y = c(8, "bold"),
           font.tickslab = 8,
           conf.int = TRUE)
#Get the Cox model predictions:
R1coxmodel <-  coxph(Surv(FUP, Progressed) ~ R1,
                   data = ASResults)
R1_HR = round(coef(R1coxmodel), 2)
R1_HRConf = confint(R1coxmodel)

R4coxmodel <-  coxph(Surv(FUP, Progressed) ~ R4,
                   data = ASResults)
R4_HR = round(coef(R4coxmodel), 2)
R4_HRConf = confint(R4coxmodel)
```

<!--- This prints the bibliography and affiliations as a single floating figure to allow for appropriate vertical spacing -->
\vspace{-15mm}
\begin{figure}
\printbibliography 
\vspace{0mm}
\centering
\includegraphics[clip=true,width = 38cm]{Affiliations}
\end{figure}
\vspace{-60mm}
<!--- Add the footnote and the QR code to the GitHub repo, with the raisebox argument adjusting the image vertical centering -->
\blfootnote{\newline All analyses presented were undertaken in R, with the poster produced through the `posterdown` package. All of the code \& data needed to reproduce this poster exactly as seen can be found at \textbf{https://github.com/Shedimus/EACR-Tracking-Cancer}\raisebox{-.4\height}{\includegraphics[clip=true,width = 4cm]{GitHubQR}}}