---
#----ESSENTIALS----#
title: '\fontfamily{phv}\selectfont Predicting outcome in prostate cancer patients using a multi-signature risk classifier, derived from urinary extracellular vesicles'
author: "Shea P. Connell^1^, Marcel Hanna^1^, Frank McCarthy^2^, Rachel Hurst^1^, Helen Curley^1^, Martyn Webb^1^, Helen Walker^3^, Rob Mills^3^, Richard Y. Ball^3^, Martin G. Sanda^4^, Kathryn L. Pellegrini^4^, Dattatraya Patil^4^, Antoinette S. Perry^5^, Jack Schalken^6^, Hardev Pandha^7^, Hayley Whitaker^8^, Nening Dennis^2^, Christine Stuttle^2^, Ian G. Mills^9, 10, 11^, Ingrid Guldvik^10^, The Movember Urine Biomarker Consortium, Chris Parker^12, 13^, Jeremy Clark^1^, Daniel S. Brewer^1, 13^, Colin S. Cooper^1^"
affiliation: "^1^ Norwich Medical School, University of East Anglia, UK; ^2^ The Institute of Cancer Research, UK; ^3^ Norfolk and Norwich University Hospitals NHS Foundation Trust, UK; ^4^ Department of Urology, Winship Cancer Institute, Emory University School of Medicine, USA; ^5^ Cancer Biology and Therapeutics Laboratory, School of Biology and Environmental Science, Conway Institute, University College Dublin, Ireland; ^6^ Nijmegen Medical Centre, Radboud University Medical Centre, The Netherlands; ^7^ Faculty of Health and Medical Sciences, The University of Surrey, UK; ^8^ Molecular Diagnostics and Therapeutics Group, University College London, UK; ^9^ School of Medicine, Dentistry and Biomedical Sciences, Institute for Health Sciences, Centre for Cancer Research and Cell Biology, Queen's University Belfast, UK; ^10^ Centre for Molecular Medicine, University of Oslo, Norway; ^11^ Nuffield Department of Surgical Sciences, University of Oxford, UK; ^12^ The Royal Marsden Hospital, Sutton, UK; ^13^ The Earlham Institute, Norwich Research Park, UK"
#----FONT STYLE----#
font_size: "30pt" #sets '\normalsize' (sizes of the title and sections headings will adjust based on it)
font_family: "palatino" # choose from typical latex fonts (example: "palatino")

#----COLOURS----#

#Background/ Line Colours
title_bgcol: "690F8C"  #Hex colour for background of poster title section (default is blue)
poster_bgcol: "ffffff" #colour of the poster main background
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
columnline_col: "000000" #colour 
header_bgcol: "008080"

#Text Colours
title_textcol: "ffffff" #colour of title text
author_textcol: "DCDCDC" # Colour of author text
affiliation_textcol: "A1A1A1" # Colour of affiliation text
body_textcol: "000000" # Colour of main text in the body of poster
header_textcol: "CC0000" # Colour of the poster header titles
footnote_textcol: "FFFFFF" # Colour of footnote text if used

#----POSTER SIZE & COLUMN FORMATTING----#
poster_height: "140cm" # height in inches of poster
poster_width: "90cm" # width in inches of poster
column_numbers: 2 # Number of columns that the poster has
column_margins: "50mm" # Margin spacing for columns

#----OTHER----#
bibliography: MyLibrary # name of the .bib file used for referencing
output: posterdown::posterdown_pdf
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis',fig.height = 3.5, fig.width = 4.5, out.width = "100%", warning = FALSE, fig.align = 'center', message = FALSE, dev='cairo_pdf')
library(readr)
library(kableExtra)
library(survival)
library(survminer)
library(pROC)
```
\definecolor{bordercolour}{HTML}{DCDCDC}
\definecolor{titleColour}{HTML}{252525}
\mdfsetup{frametitlealignment=\center,
          frametitlefont=\fontfamily{ppl}\selectfont,
          frametitlefont=\huge,
          frametitlebackgroundcolor=bordercolour,
          frametitlefontcolor = titleColour,
          linecolor=black,
          linewidth=5pt}
\definecolor{myframecolour}{HTML}{690F8C}
\begin{mdframed}[backgroundcolor = myframecolour,
                  font = \fontfamily{ppl}\selectfont,
                  fontcolor = white, 
                  roundcorner = 20pt, 
                  innerleftmargin = 30pt,
                  innerrightmargin = 30pt,
                  frametitle={\textbf{Quick Read:}}]
\large
\textbf{Question:} Can a non-invasive single-sample urine test reveal \textbf{both diagnostic and prognostic information} about 537 prostate cancer patients, utilising extracellular vesicle-derived RNA expression patterns? \\\\
\textbf{Findings:} A robust, four-risk-signature model identified two groups with differing rates of treatment intervention in active surveillance use (\textbf{High Risk HR = 3.7, Low Risk HR = -7.0}) and predicted initial biopsy outcome (\textbf{AUC = 0.81})\\\\
\textbf{Impact:} Clinical implementation of this model has the potential to avoid the unnecessary initial biopsy of men and the repeated, invasive follow-up of men on active surveillance with indolent prostate cancer could be drastically reduced, or provide a means for exiting surveillance altogether.\\
\end{mdframed}
\vspace{-40mm}

# Introduction

Prostate cancer is estimated to be diagnosed in 12% of men, the most commonly diagnosed cancer in men in the UK[@CancerResearchUK2015]. However, 5-year survival of men with prostate cancer is very high, and treatment of indolent, clinically "irrelevant" disease has significant side-effects, including incontinence and impotence.

Men diagnosed with lower risk cancer are often placed on active surveillance (AS) programmes, which have proven successful in improving survival rates. Regardless, there is still no formal method for patients on AS to exit these programmes. 

Urine represents a suitable medium for non-invasively sampling the prostate[@McKiernan2016b;@Tomlins2016a;@Donovan2015]. In this vain, we have developed a risk prediction model based on NanoString quantified RNA expression in urinary extracellular vesicles, with the aim to discriminate accurately between men with, and without, clinically significant cancers. 
\vspace{-30mm} 

# The Movember Cohort:

  **Patients:** 537 men with and without histologically proven cancer with a PSA < 100 ng/mL, collected from urology clinics in the UK, Ireland and USA between 2009 and 2015. 
  
  **Samples:** NanoString extracellular vesicular RNA expression profiles of 167 gene-probes, derived from first-catch post-digital rectal examination urine. 
  
  **AS Eligibility:** Age 50–80, clinical stage T1/T2, PSA < 15 ng/mL, Gs <7 (Gs < 4+3 if age >65), and <50% percent positive biopsy cores.  
  
  **Progression:** PSA velocity >1 ng/mL per year or adverse histology on repeat biopsy, defined as primary G >3 or >50% biopsy cores positive for cancer 

\vspace{-25mm}

# Methods:
A continuation-ratio LASSO model was trained on 67% of the data, using D'Amico risk status as an ordinal variable, producing four risk signatures for predicting the probability of normal tissue (R1), D’Amico Low-risk (R2), Intermediate-risk (R3), and High-risk (R4) PCa in a given sample. 

This model was applied to the test dataset and to an AS sub-cohort (n = 87) for diagnostic and prognostic evaluation, respectively. ROC-AUC analysis was used to test predicting biopsy outcome, and survival analyses used to prognosticate the progression of patient in the AS cohort. 
\vspace{-35mm}

# Results
R4 was able to determine a clinically significant biopsy outcome of D'Amico Intermediate- or High-Risk from a single urine sample, with an **AUC = 0.813 (95% CI: 0.779 - 0.847, Figure 1)**. 

A robust R1 threshold dichotomised AS patients in to groups at low and high risk of progression. The two groups had a large difference in time to progression: 60% progression within 5 years of urine sample collection in the poor prognosis group compared to 10% in the good prognosis group ($p$ < 0.001, log-rank test, Figure 2). 

```{r, fig.cap='ROC-AUC plot of the ability of R4 to predict the presence of D\'Amico Intermediate or High risk cancer on initial biopsy.', cache=TRUE, fig.asp = 0.8, fig.width=5, out.extra='trim={0 5mm 0 10mm},clip'}
#Read in the pre-cleaned data:
Results <- readRDS("data/Results.RDS") 
par(bty = "n")
#Create the ROC object for R4:
R4roc <- roc(response = Results$ClinSig,
             predictor = Results$R4,
             ci = TRUE, 
             ci.method = "bootstrap",
             # arguments for plot
             plot=TRUE,
             print.auc=TRUE, 
             show.thres=TRUE,
             print.auc.x = 0.5,
             print.auc.y = 0.05,
             print.auc.cex = 0.85,
             asp = 0.75,
             cex.axis = 0.75,
             xlab = "",
             ylab = "",
             identity.col = "black"
             
)
# Compute and plot the 95% CI:
sens.ci <- ci.se(R4roc,
                 specificities = seq(0, 1, 0.05))
RedCol <- adjustcolor("red2", alpha.f = 0.4) #Custom colour allows for alpha
plot(sens.ci, 
     type ="shape",
     lwd = 0.1,
     col = RedCol,
     alpha = 0.75)
# need to re-draw R4roc over the CI
plot(R4roc, add=TRUE, col = "red2")
mtext(side=1, line=3, "Specificity")
mtext(side=2, line=2, "Sensitivity")
```
\vspace{-10mm}
Cox proportional hazards models showed significant differences in clinical outcome based on R1 or R4 of patients. The hazard ratio associated with R4 was **3.71 (95% CI: 1.53 to 5.89)**, with the **HR of R1 =-7.03 (95% CI: -12.29 to -1.77)**.


```{r, fig.asp=0.75, fig.cap='Kaplan-Meier plot of time to disease progression in months from initial urine collection. Colours indicate the dichotomised model thresholds, Green – High R1, Red – Low R1. Numbers above the x-axis indicate the number of patients in each group at risk at the given time intervals.', out.extra='trim={0 2.5mm 0 2.5mm},clip'}
#Read in the pre-cleaned data:
ASResults <- readRDS("data/ASResults.RDS") 
#Optimise the threshold to find the best binary threshold that keeps at least 30% of patients in the smallest group
OptimalThresh <- surv_categorize(surv_cutpoint(data = ASResults,
                                                    time = "FUP",
                                                    event = "Progressed",
                                                    variables = "R1",
                                                    minprop = 0.30),
                                      labels = c("Low R1", "High R1"))
#Build the survival object 
SurvOptim <- survfit(Surv(FUP, Progressed) ~ R1, data = OptimalThresh)

#Construct the plot (using SurvMiner)
ggsurvplot(SurvOptim, 
           risk.table = TRUE,
           palette =  c("ForestGreen", "Red2"),
           legend = "none",
           risk.table.title =  "",
           risk.table.pos = "in",
           xlab = "Time (months)",
           ylab = "Proportion free of progresison",
           tables.height = 0.25,
           font.x = c(8, "bold"),
           font.y = c(8, "bold"),
           conf.int = TRUE)
#Get the Cox model predictions:
R1coxmodel <-  coxph(Surv(FUP, Progressed) ~ R1,
                   data = ASResults)
R1_HR = round(coef(R1coxmodel), 2)
R1_HRConf = confint(R1coxmodel)

R4coxmodel <-  coxph(Surv(FUP, Progressed) ~ R4,
                   data = ASResults)
R4_HR = round(coef(R4coxmodel), 2)
R4_HRConf = confint(R4coxmodel)
```

<!--- Here you can set the size of the citation text as well as remove the "References" section if you choose not to have one for some reason :) -->
\vspace{-45mm}
\footnotesize\printbibliography 

\vspace{5mm}
\begin{figure}
\includegraphics[width = 10cm,right]{UEA_NEW_BRAND_Cyan_N_A}
\end{figure}
